\pkg{Rgbp}: An R Package for Hierarchical Modeling and Method Checking

### 2013-05-29 jss/article.tex
uhhh more changes

\section[Multilevel Structure]{Multilevel Structure}
Our multilevel model, also called a conditionally independent hierarchical model \cite{1989}, is a very powerful tool for exploring the hierarchical structure in data. For example, we can think about a district-level hierarchy (bigger population) for 8 schools, the state-level hierarchy for 31 hospitals, and the position-level hierarchy for 18 baseball players. \code{gbp}, one of functions in \pkg{Rgbp}, fits such a hierarchical model whose first-level hierarchy has a distribution of observed data and second-level (bigger population hierarchy) has a conjugate prior distribution on the first-level parameter. Users can determine one of three types of multilevel models, such as Normal-Normal, Poisson-Gamma, and Binomial-Beta, based on their data sets. 
\\

\subsection[Normal-Normal]{Normal-Normal}
\code{gbp} can construct a two-level Normal-Normal hierarchical model on the 8 school data. For reference,  $V_{j}~(\equiv \sigma^{2}/n_{j})$ below is assumed to be known or to be accurately estimated, and subscript \emph{j} indicates \emph{j}-th school in the data set.
\begin{equation}
y_{j}\vert \mu_{j} \stackrel{ind}{\sim}\textrm{Normal}(\mu_{j}, V_{j}),
\end{equation}
\begin{equation}
\mu_{j}\vert \beta, A\stackrel{ind}{\sim}\textrm{Normal}(\mu_{0j}, A),
\end{equation}

where $\mu_{0j} =x^{T}_{j}\beta,~j=1, \ldots, 8$, $\beta$ and $A$ are unknown, $x_{j}$ is $j$-th school's covariate vector ($m\times 1$), and $m$ is the number of regression coefficients. Note that if there is no covariate then $x_{j}=1$ for an intercept term ($m=1$) and so $\mu_{0j}=\mu_{0}=\beta_{0}$ for all $j$, resulting in an exchangeable prior distribution. For reference, a parameter with a zero subscript, such as $\mu_{0j}$, represents a mean parameter of the prior (second-level) distribution, \emph{i.e.}, a prior mean. Also,  based on this conjugate prior distribution, it is easy to derive the corresponding posterior distribution, which is the same as the product of two distributions, (1) and (2), up to a normalizing constant.
\begin{equation}
\mu_{j}\vert \textbf{y}, \beta, A \stackrel{ind}{\sim}\textrm{Normal}(~(1-B_{j})y_{j} + B_{j}\mu_{0j},~(1-B_{j})V_{j}),
\end{equation}
where $B_{j}\equiv V_{j}/(V_{j} + A),~j=1, \ldots, 8$, are called shrinkages.

\subsection[Poisson-Gamma]{Poisson-Gamma}
\code{gbp} is also able to build a Poisson-Gamma multilevel model on the 31 hospital data. Note that a constant, $1/r$, multiplied to the Gamma distribution below is a scale and a square bracket below indicates [mean, variance] of distribution. And for notational consistency, let's define $y_{j}\equiv z_{j} / n_{j}$ for all $j$.
\begin{equation}
z_{j}\vert \lambda_{j} \stackrel{ind}{\sim}\textrm{Poisson}(n_{j}\lambda_{j}),
\end{equation}
\begin{equation}
\lambda_{j}\vert \beta, r\stackrel{ind}{\sim}\frac{1}{r}\textrm{Gamma}(\lambda_{0j}r)\sim \textrm{Gamma}[\lambda_{0j}, ~\frac{\lambda_{0j}}{r}],
\end{equation}

where $\log(\lambda_{0j}) =x_{j}'\beta$, $j=1, \ldots, 31$, with two unknown hyper-parameters, $r$ and $\beta$. Immediate posterior distribution of this Poisson-Gamma model is
\begin{equation}
\lambda_{j}\vert \textbf{z}, \beta, r \stackrel{ind}{\sim}\frac{1}{r + n_{j}}\textrm{Gamma}(r\lambda_{0j} + n_{j}y_{j})\sim\textrm{Gamma}[\lambda^{\ast}_{j},~\frac{\lambda^{\ast}_{j}}{r+n_{j}}],
\end{equation}

where $\lambda^{\ast}_{j} \equiv (1-B_{j})y_{j} + B_{j}\lambda_{0j}$,  $B_{j}\equiv r / (r+n_{j})$, and $y_{j}\equiv z_{j} / n_{j}$, $j=1, \ldots, 31$.

\subsection[Binomial-Beta]{Binomial-Beta}
Binomial-Beta hierarchical model is the last model that \code{gbp} can fit. Again, a square bracket below indicates [mean, variance] of distribution.
\begin{equation}
z_{j} \vert p_{j}\stackrel{ind}{\sim}\textrm{Binomial}(n_{j}, ~p_{j}),
\end{equation}
\begin{equation}
p_{j} \vert \beta, r\stackrel{ind}{\sim}\textrm{Beta}(rp_{0j},~ r(1-p_{0j}))\sim \textrm{Beta}[p_{0j}, ~\frac{p_{0j}(1-p_{0j})}{r + 1}],
\end{equation}

where $\log(\frac{p_{0j}}{1-p_{0j}}) =x_{j}'\beta, ~j=1, \ldots, 18$. Note that $r$ and $\beta$ are two unknown hyper-parameters. Then, corresponding posterior distribution is
\begin{equation}
p_{j}\vert \textbf{z}, \beta, r \stackrel{ind}{\sim}\textrm{Beta}(rp_{0j}+n_{j}y_{j},~r(1-p_{0j})+n_{j}(1-y_{j}))\sim\textrm{Beta}\bigg[p^{\ast}_{j},~ \frac{p^{\ast}_{j}(1-p^{\ast}_{j})}{r+n_{j}+1}\bigg],
\end{equation}
where $p^{\ast}_{j}\equiv(1-B_{j})y_{j}+B_{j}p_{0j}$, $B_{j}\equiv r/ (r+n_{j})$, and $y_{j}\equiv z_{j} / n_{j}$, $j=1,\ldots,18$.

\subsection[Hyper-prior Distribution]{Hyper-prior Distribution}
Hyper-prior distribution indicates a distribution of the second-level parameters, which plays an important role in deriving a full posterior distribution of all the parameters. \code{gbp} sets non-informative distributions on second-level parameters to let the data speak more about their estimation.
\begin{equation}
\beta \sim \textrm{Uniform on}~ \mathbf{R}^{m}~~\textrm{and}~~A \sim \textrm{Uniform}(0, \infty) ~~(\textrm{or} ~\frac{1}{r}\sim \textrm{Uniform}(0, \infty)),
\end{equation}

where $m$ is the number of regression coefficients. As for $\beta$, it is reasonable choice to take flat (non-informative) distribution because information about the location gets plentiful as the number of groups increases. The next flat prior distribution of the second-level variance component $A$ (or $1/r$) was chosen to bring a good repeated sampling property and for a posterior propriety under moderate conditions.
\\

@@

\subsection[Shrinkage Estimation]{Shrinkage Estimation}
Estimating shrinkage is a key part of this hierarchical modeling. As we can see in (3), (6), and (9), the posterior means are a linear function of shrinkage and the posterior variances are also a linear (Gaussian), quadratic (Poisson), or cubic (Binomial) function of shrinkage. It implies that we cannot obtain $E(\mu_{j}\vert \textbf{y})$ and $Var(\mu_{j}\vert \textbf{y})$ without appropriate shrinkage estimation.

\subsection[ADM]{Adjustment for Density Maximization}
When it comes to estimating a shrinkage, we can notice that it is a function of the second-level variance component, \emph{i.e.}, $B_{j}\equiv V_{j}/(V_{j}+A)=B_{j}(A)$ for Gaussian and $B_{j}\equiv r/(r+n_{j})=B_{j}(r)$ for Poisson and Binomial models.
\\

In this case, the most common way is to obtain an MLE of the variance component with its asymptotic Normality and then to use Delta method for asymptotic Normal distribution of shrinkage, \emph{i.e.}, $\hat{B}_{j, MLE}=B_{j}(\hat{A}_{MLE})$. But, is the Normal approximation a good approximation for shrinkage that takes on a value between 0 and 1? If it is, why do we fit a logistic regression, such as logit($p$) = $x^{T}\beta$, instead of fitting a linear regression, $p = x^{T}\beta+\epsilon$?
\\

Here the adjustment for density maximization \cite{2011}, called ADM, comes. It assumes the Beta distribution for shrinkage because this distribution is also defined between 0 and 1. In the end, ADM  estimates its posterior moments, \emph{i.e.}, $E(B_{j}\vert\textrm{data})$ and $Var(B_{j}\vert\textrm{data})$, without any trouble that MLE can cause. Please refer to \cite{2011} for additional advantages of ADM.
\\

Once we estimate these two moments of shrinkage, we can also estimate the posterior moments given only data. Taking the Normal model as an example,  $E(\mu_{j}\vert \textbf{y})=E(E(\mu_{j}\vert \textbf{y}, r, A)\vert \textbf{y})$ by Adam's law and $Var(\mu_{j}\vert \textbf{y})=E(Var(\mu_{j}\vert \textbf{y}, r, A)\vert \textbf{y})+Var(E(\mu_{j}\vert \textbf{y}, r, A)\vert \textbf{y})$ by Eve's law. Note that both $E(\mu_{j}\vert \textbf{y}, r, A)$ and $Var(\mu_{j}\vert \textbf{y}, r, A)$ are functions of shrinkage (see (3)) and Adam's and Eve's laws are taking conditional expectation and variance over them given data, $\textbf{y}$. Since we have already estimated $E(B_{j}\vert\textbf{y})$ and $Var(B_{j}\vert\textbf{y})$ via ADM, finally we can estimate $E(\mu_{j}\vert \textbf{y})$ and $Var(\mu_{j}\vert \textbf{y})$.

\subsection[Approximation to Posterior Distribution by Moment Matching]{Approximation to Posterior Distribution via Matching Moments}
After estimating two posterior moments, $E(p_{j}\vert \textbf{z})$ and $Var(p_{j}\vert \textbf{z})$ for the Binomial model, \code{gbp} reasonably approximates a posterior distribution given data, \emph{i.e.}, $p_{j}\vert \textbf{z}$ (or $\mu_{j}\vert \textbf{y}$, $\lambda_{j}\vert \textbf{z}$), by matching these two moments with its parameters. To be specific, we assumed $p_{j}\vert \textbf{z}$ had another Beta($a_{1}$, $a_{0}$) distribution and matched two estimated moments, $E(p_{j}\vert \textbf{z})$ and $Var(p_{j}\vert \textbf{z})$, with two parameters, $a_{1}$ and $a_{0}$, of this assumed Beta distribution.

\section[Method Checking]{Method Checking}
Like the two sides of the same coin, checking a statistical model always comes with fitting a model. If a fitted model cannot pass a checking process, we usually go back to the fitting process and come back to the checking process iteratively. In this sense, checking a fitted model is an interactive procedure for the model justification.
\\

There are two kinds of model justification process; one is a model check and the other is a method check. The model check is for the justification of a hierarchical modeling on a specific data set. One possible question is, ``Can this data set benefit from such a multilevel modeling?'' Christiansen and Morris (1996) answered this question by using a mixture model, $z_{j}\vert \beta, r\sim\textrm{Negative-Binomial}$, on Poisson data to justify the second-level hierarchy. They found that their data had more variation than expected of the first-level Poisson distribution and Poisson hierarchical model could successfully account for such additional variation.
\\

@@

\section[Examples]{Examples}
\subsection[Known Second-level Mean]{31 Hospitals: Known Second-level Mean}
Suppose you live in the New York state (NY) and have been suffering from severe coronary heart disease (hopefully not). If you are supposed to receive the coronary artery bypass graft (CABG) surgery soon, you might want to find the most famous hospital for dealing with such a surgery. On top of that, if you can figure out each hospital's ability to handle this surgery, it will be useful for your decision in choosing a hospital.
\\

For this purpose, you gathered data of 31 hospitals in NY composed of the number of deaths within a month of CABG surgeries and total number of patients receiving CABG surgeries in each hospital. In addition, while one looks for such information, suppose one knows that the state-level death rate per exposure of this surgery is 0.020. 
\\

The multilevel modeling that assumes a bigger population-level hierarchy will be insightful in this problem. Here, we presume a state-level (NY) hierarchy governing the true death rates of CABG surgery of all the hospitals in NY. This perspective is to view the true death rates of those 31 hospitals as sampled from the state-level population distribution whose mean is 0.020. For reference, a model check can be useful for evaluating the validity of such a view point (Christiansen and Morris, 1996). 
\\

@@

         obs.mean    n prior.mean shrinkage low.intv post.mean upp.intv post.sd
1           0.045   67       0.02     0.834    0.011     0.024    0.042   0.008
2           0.029   68       0.02     0.832    0.010     0.022    0.038   0.007
3           0.024  210       0.02     0.616    0.011     0.021    0.035   0.006
4           0.043  256       0.02     0.568    0.017     0.030    0.046   0.008
5           0.033  269       0.02     0.556    0.015     0.026    0.041   0.007
6           0.044  274       0.02     0.551    0.018     0.031    0.047   0.008
7           0.043  278       0.02     0.548    0.018     0.030    0.047   0.007
8           0.014  295       0.02     0.533    0.008     0.017    0.029   0.005
9           0.029  347       0.02     0.492    0.014     0.024    0.038   0.006
10          0.037  349       0.02     0.491    0.017     0.029    0.043   0.007
11          0.039  358       0.02     0.484    0.018     0.030    0.045   0.007
12          0.018  396       0.02     0.459    0.010     0.019    0.030   0.005
13          0.028  431       0.02     0.438    0.015     0.024    0.037   0.006
14          0.025  441       0.02     0.433    0.013     0.023    0.035   0.005
15          0.027  477       0.02     0.414    0.015     0.024    0.036   0.006
16          0.045  484       0.02     0.410    0.023     0.035    0.050   0.007
17          0.030  494       0.02     0.405    0.016     0.026    0.039   0.006
18          0.022  501       0.02     0.402    0.012     0.021    0.032   0.005
19          0.028  505       0.02     0.400    0.015     0.025    0.036   0.005
20          0.020  540       0.02     0.384    0.012     0.020    0.031   0.005
21          0.028  563       0.02     0.374    0.016     0.025    0.037   0.005
22          0.024  593       0.02     0.362    0.014     0.022    0.033   0.005
23          0.015  602       0.02     0.358    0.010     0.017    0.026   0.004
24          0.024  629       0.02     0.348    0.014     0.023    0.033   0.005
25          0.020  636       0.02     0.346    0.012     0.020    0.030   0.005
26          0.048  729       0.02     0.316    0.027     0.039    0.053   0.007
27          0.031  849       0.02     0.284    0.019     0.028    0.038   0.005
28          0.027  914       0.02     0.269    0.017     0.025    0.035   0.005
29          0.021  940       0.02     0.264    0.014     0.021    0.030   0.004
30          0.029 1193       0.02     0.220    0.020     0.027    0.036   0.004
31          0.020 1340       0.02     0.201    0.014     0.020    0.027   0.003
colMeans    0.029  517       0.02     0.438    0.015     0.025    0.037   0.006
\end{CodeOutput}
\end{CodeChunk}
For reference, we need to type `\code{R> print(p, sort = FALSE)}' instead of `\code{R> p}' in order to list hospitals by the order of data input in the above output. `\code{R> p}' automatically sorts the output by the increasing order of caseload ($n_{j}$). 
\\

The output contains information about sample mean (\code{obs.mean}), caseload (\code{n}), known prior mean ($\lambda_{0}$), shrinkage estimate ($\hat{B}_{j}$), lower interval, posterior mean ($E(\lambda_{j}\vert \textbf{z})$), upper interval, and standard deviation of posterior distribution ($sd(\lambda_{j}\vert \textbf{z})$).
\\

As we can see in (6), the posterior mean, $(1-B_{j})y_{j} + B_{j}\lambda_{0}$, is a linear function of shrinkage, $B_{j}\equiv r / (r + n_{j})$, locating between the sample mean and prior mean ($\lambda_{0}=0.02$). It makes sense because $r$ can be interpreted as the amount of prior information and $n_{j}$ as the amount of observed information. If the second level has more information than the first level, then the sample mean shrinks towards the prior mean more than 50\%. This point is clear in the above output because, as caseload increases, shrinkage decreases, depending less on the prior distribution.
\\

@@

  post.mode.alpha post.sd.alpha
1          -5.818         0.411
\end{CodeOutput}
\end{CodeChunk}
Since estimated $\alpha$ is -5.818, we can easily calculate $\hat{r}=\textrm{exp}(5.818)=336$, which is a good indicator of how valuable and informative the hypothetical second-level hierarchy is. As mentioned before, $r$ can be interpreted as the amount of prior information, although we do not know its true value. So, \code{gbp} asked data how much credit they wanted to give to this hypothetical second-level hierarchy. $\hat{r}$ = 336 is the answer from the data. It means that observed sample means of hospitals whose caseloads are less than 336 will shrink toward the prior mean (0.020) more than 50\%. 
\\

For example, the shrinkage estimate of the first hospital ($\hat{B}_{1}= 0.834$) was calculated by 336 / (336 + 67), where 67 is its caseload ($n_{1}$), and its posterior mean is $(1-0.834)*0.045 + 0.834 * 0.020=0.024$. As for this hospital, using more information from the prior distribution is an appropriate choice because its observed amount of information (67) is far less than the amount of state-level information (336).
\\

@@

The regression towards the mean (RTTM) is obvious in the left-side graph; the observed sample means, empty dots on the upper line, are shrinking towards the known second-level mean (a blue vertical line at 0.02) to the different extents. Note that some hospitals' ranks have changed by shrinking much harder towards 0.02 than others. For example, the empty square symbol at the crossing point of the two left-most lines (8th and 23rd hospitals on the list above) indicates that seemingly safest hospital among 31 hospitals in terms of the observed sample mean was not actually safer than the second safest hospital. Without this hierarchical modeling, we might have made a wrong decision in choosing a hospital.
\\

>>>>

\section[Multilevel Structure]{Multilevel Structure} \label{multilevel}
A two-level or multilevel model, also called a conditionally independent hierarchical model \cite{1989}, is a very powerful tool for exploring the hierarchical structure in data. For example, in the 8 schools dataset, we can imagine that there exists a district-level hierarchy (bigger population) for 8 schools, the state-level hierarchy for 31 hospitals, and the position-level hierarchy for 18 baseball players. \code{gbp}, one of the functions in \pkg{Rgbp}, fits such a hierarchical model whose first-level hierarchy has a distribution of observed data and second-level (bigger population hierarchy) has a conjugate prior distribution on the first-level parameter. The \code{gbp} function allows users to choose one of three types of multilevel models, such as Normal-Normal, Poisson-Gamma, and Binomial-Beta, based on their data sets.

\subsection[Normal-Normal]{Normal-Normal}
The following is a general Normal-Normal hierarchical model. For reference,  $V_{j}~(\equiv \sigma^{2}/n_{j})$ below is assumed to be known or to be accurately estimated, and subscript \emph{j} indicates \emph{j}-th group in the data set.
\begin{equation}
y_{j}\vert \mu_{j} \stackrel{ind}{\sim}\textrm{Normal}(\mu_{j}, V_{j}),
\end{equation}
\begin{equation}
\mu_{j}\vert \beta, A\stackrel{ind}{\sim}\textrm{Normal}(\mu_{0j}, A),
\end{equation}

where $~j=1, \ldots k$, $\mu_{0j} = \mathbf{x}^{T}_{j} \mathbf{\beta}$, $x_{j}$ is $j$-th group's covariate vector ($m\times 1$), $m$ is the number of regression coefficients and $\mathbf{\beta}$ and $A$ are unknown. Note that if there is no covariate then $x_{j}=1$ for an intercept term ($m=1$) and so $\mu_{0j}=\mu_{0}=\beta_{0}$ for all $j$, resulting in an exchangeable prior distribution. For reference, a parameter with a zero subscript, such as $\mu_{0j}$, represents a mean parameter of the prior (second-level) distribution, \emph{i.e.}, a prior mean. Based on this conjugate prior distribution it is easy to derive the corresponding posterior distribution
\begin{equation} \label{normalpost}
\mu_{j}\vert \textbf{y}, \beta, A \stackrel{ind}{\sim}\textrm{Normal}(~(1-B_{j})y_{j} + B_{j}\mu_{0j},~(1-B_{j})V_{j}),
\end{equation}
where $B_{j}\equiv V_{j}/(V_{j} + A),~j=1, \ldots, k$, are called shrinkages.

\subsection[Poisson-Gamma]{Poisson-Gamma}
\code{gbp} is also able to build a Poisson-Gamma multilevel model. Note that a constant, $1/r$, multiplied to the Gamma distribution below is a scale and a square bracket below indicates [mean, variance] of distribution. And for notational consistency, let's define $y_{j}\equiv z_{j} / n_{j}$ for all $j$.
\begin{equation}
z_{j}\vert \lambda_{j} \stackrel{ind}{\sim}\textrm{Poisson}(n_{j}\lambda_{j}),
\end{equation}
\begin{equation}
\lambda_{j}\vert \beta, r\stackrel{ind}{\sim}\frac{1}{r}\textrm{Gamma}(\lambda_{0j}r)\sim \textrm{Gamma} \left[\lambda_{0j}, ~\frac{\lambda_{0j}}{r} \right],
\end{equation}

where $\log(\lambda_{0j}) =x_{j}'\beta$, $j=1, \ldots, k$, with two unknown hyper-parameters, $r$ and $\beta$. The posterior distribution of this Poisson-Gamma model is
\begin{equation} \label{gammapost}
\lambda_{j}\vert \textbf{z}, \beta, r \stackrel{ind}{\sim}\frac{1}{r + n_{j}}\textrm{Gamma}(r\lambda_{0j} + n_{j}y_{j})\sim\textrm{Gamma} \left[\lambda^{\ast}_{j},~\frac{\lambda^{\ast}_{j}}{r+n_{j}} \right],
\end{equation}

where $\lambda^{\ast}_{j} \equiv (1-B_{j})y_{j} + B_{j}\lambda_{0j}$,  $B_{j}\equiv r / (r+n_{j})$, and $y_{j}\equiv z_{j} / n_{j}$, $j=1, \ldots, k$.

\subsection[Binomial-Beta]{Binomial-Beta}
Binomial-Beta hierarchical model is the last model that \code{gbp} can fit. Again, a square bracket below indicates [mean, variance] of distribution.
\begin{equation}
z_{j} \vert p_{j}\stackrel{ind}{\sim}\textrm{Binomial}(n_{j}, ~p_{j}),
\end{equation}
\begin{equation}
p_{j} \vert \beta, r\stackrel{ind}{\sim}\textrm{Beta}(rp_{0j},~ r(1-p_{0j}))\sim \textrm{Beta} \left[p_{0j}, ~\frac{p_{0j}(1-p_{0j})}{r + 1} \right],
\end{equation}

where $\log(\frac{p_{0j}}{1-p_{0j}}) =x_{j}'\beta, ~j=1, \ldots, k$. Note that $r$ and $\beta$ are two unknown hyper-parameters. Then, corresponding posterior distribution is
\begin{equation} \label{betapost}
p_{j}\vert \textbf{z}, \beta, r \stackrel{ind}{\sim}\textrm{Beta}(rp_{0j}+n_{j}y_{j},~r(1-p_{0j})+n_{j}(1-y_{j}))\sim\textrm{Beta}\bigg[p^{\ast}_{j},~ \frac{p^{\ast}_{j}(1-p^{\ast}_{j})}{r+n_{j}+1}\bigg],
\end{equation}
where $p^{\ast}_{j}\equiv(1-B_{j})y_{j}+B_{j}p_{0j}$, $B_{j}\equiv r/ (r+n_{j})$, and $y_{j}\equiv z_{j} / n_{j}$, $j=1,\ldots,k$.

\subsection[Hyper-prior Distribution]{Hyper-prior Distribution}
In our examples the hyper-prior distribution is an assumed distribution on the second-level parameters and with the goal of objectivity in mind \code{gbp} assumes non-informative hyper-prior distributions.
\begin{equation}
\beta \sim \textrm{Uniform on}~ \mathbf{R}^{m}~~\textrm{and}~~A \sim \textrm{Uniform}(0, \infty) ~~(\textrm{or} ~\frac{1}{r}\sim \textrm{Uniform}(0, \infty)),
\end{equation}

where $m$ is the number of regression coefficients. As for $\beta$, it is a reasonable choice to assume a flat (non-informative) distribution because information about the location gets plentiful as the number of groups increases. The next flat prior distribution of the second-level variance component $A$ (or $1/r$) was chosen for its good repeated sampling properties and for posterior propriety under moderate conditions.

@@

\subsection[Shrinkage Estimation]{Shrinkage Estimation}
Estimating the shrinkage factors ($B_1, \ldots B_k$) is the key estimation problem with the hierarchical models \code{gbp} assumes. As we can see in \ref{normalpost}, \ref{gammapost}, and \ref{betapost}, the posterior means are a linear function of the shrinkage factors and the posterior variances are also linear (Gaussian), quadratic (Poisson), or cubic (Binomial) functions. A natural method then to estimate $E(\mu_{j}\vert \textbf{y})$ and $Var(\mu_{j}\vert \textbf{y})$ is to first estimate the shrinkage factors.

\subsection[ADM]{Adjustment for Density Maximization}
It is noted that the shrinkage factors ($B_1, \ldots B_k$) are a function of the second-level variance component, \emph{i.e.}, $B_{j}\equiv V_{j}/(V_{j}+A)=B_{j}(A)$ for Gaussian and $B_{j}\equiv r/(r+n_{j})=B_{j}(r)$ for Poisson and Binomial models. Current popular methods of estimation via maximization (MLE or MAP) of these unknown shrinkage factors can result in estimates lying on the boundary of the parameter space. In the Normal-Normal model this situation typically arises when the variation within schools is greater than the variation between resulting in a parameter estimate of $\hat{A} = 0$. %TODO: Reference this! Maybe it's in the Lysy paper

To continue with a maximization based estimation procedure but to steer clear of aforementioned issues we make us of adjustment for density maximization (ADM) \cite{2011}. In general ADM is a procedure to obtain estimates of posterior moments via maximization if the underlying posterior can be approximated by a Pearson family. For our purposes we can approximate the posterior distribution of a shrinkage factor with the Beta distribution allowing us to finally obtain estimates the posterior moments, \emph{i.e.}, $E(B_{j}\vert\textrm{data})$ and $Var(B_{j}\vert\textrm{data})$, without any trouble that MLE can cause. Please refer to \cite{2011} for additional advantages of ADM.

Once we estimate these two moments of shrinkage, we can also estimate the posterior moments given only the data. Taking the Normal-Normal model as an example we note the following identities
\begin{eqnarray}
E(\mu_{j}\vert \textbf{y}) & = & E(E(\mu_{j}\vert \textbf{y}, r, A)\vert \textbf{y}) \\ 
Var(\mu_{j}\vert \textbf{y}) & = & E(Var(\mu_{j}\vert \textbf{y}, r, A)\vert \textbf{y})+Var(E(\mu_{j}\vert \textbf{y}, r, A)\vert \textbf{y}).
\end{eqnarray}  
Note that both $E(\mu_{j}\vert \textbf{y}, r, A)$ and $Var(\mu_{j}\vert \textbf{y}, r, A)$ are functions of the shrinkage factors (\ref{normalpost}) and since we have already can estimate $E(B_{j}\vert\textbf{y})$ and $Var(B_{j}\vert\textbf{y})$ via ADM, we can finally estimate $E(\mu_{j}\vert \textbf{y})$ and $Var(\mu_{j}\vert \textbf{y})$.

\subsection[Approximation to Posterior Distribution by Moment Matching]{Approximation to Posterior Distribution via Matching Moments}
After estimating the two posterior moments, $E(p_{j}\vert \textbf{z})$ and $Var(p_{j}\vert \textbf{z})$, \code{gbp} reasonably approximates a posterior distribution of the mean effects given the data by assuming a reasonable distribution and matching moments. The Binomial-Beta model we approximate $p_{j}\vert \textbf{z}$ with a Beta distribution and for the Poisson-Gamma model we approximate $\lambda_{j}\vert \textbf{z}$ with a Gamma distribution. Finally for the Normal-Normal model we actually estimate the first three moments and approximate $\mu_{j}\vert \textbf{y}$ with a skewed-t distribution. %TODO: should I put down the form of the skewed-t or just put a reference to it?

%not sure we need this
%To be specific, we assumed $p_{j}\vert \textbf{z}$ had another Beta($a_{1}$, $a_{0}$) distribution and matched two estimated moments, $E(p_{j}\vert \textbf{z})$ and $Var(p_{j}\vert \textbf{z})$, with two parameters, $a_{1}$ and $a_{0}$, of this assumed Beta distribution.

%TODO: Resume editing from here
\section[Method Checking]{Method Checking}
Like the two sides of the same coin, checking a statistical model always comes with fitting a model. If a fitted model cannot pass a validation or checking process, we usually go back and forth from estimation and checking steps iteratively. In this sense, checking a fitted model is an interactive procedure for the model justification.

There are two kinds of model justification processes; one is a model check and the other is a method check. The model check is for the justification of the model \ref{multilevel} on a specific data set. One possible question is, ``Can this data set benefit from such a multilevel modeling?'' \cite{1997} answered this question by using a mixture model, $z_{j}\vert \beta, r\sim\textrm{Negative-Binomial}$, on Poisson data to justify the second-level hierarchy. They found that their data had more variation than expected of the first-level Poisson distribution and Poisson hierarchical model could successfully account for such additional variation.

@@

\section[Examples]{Examples}
\subsection[Known Second-level Mean]{31 Hospitals: Known Second-level Mean}
Suppose you live in the New York state (NY) and have been suffering from severe coronary heart disease (hopefully not). If you are supposed to receive the coronary artery bypass graft (CABG) surgery soon, you might want to find the most famous hospital for dealing with such a surgery. On top of that, if you can figure out each hospital's ability to handle this surgery, it will be useful for your decision in choosing a hospital.

For this purpose, you gathered data of 31 hospitals in NY composed of the number of deaths within a month of CABG surgeries and total number of patients receiving CABG surgeries in each hospital. In addition, while one looks for such information, suppose one knows that the state-level death rate per exposure of this surgery is 0.020.

The multilevel modeling that assumes a bigger population-level hierarchy will be insightful in this problem. Here, we presume a state-level (NY) hierarchy governing the true death rates of CABG surgery of all the hospitals in NY. This perspective is to view the true death rates of those 31 hospitals as sampled from the state-level population distribution whose mean is 0.020. For reference, a model check can be useful for evaluating the validity of such a view point (Christiansen and Morris, 1996).

@@

         obs.mean    n prior.mean shrinkage low.intv post.mean upp.intv post.sd
1           0.045   67       0.02     0.834    0.011     0.024    0.042   0.008
2           0.029   68       0.02     0.832    0.010     0.022    0.038   0.007
3           0.024  210       0.02     0.616    0.011     0.021    0.035   0.006
4           0.043  256       0.02     0.568    0.017     0.030    0.046   0.008
5           0.033  269       0.02     0.556    0.015     0.026    0.041   0.007
6           0.044  274       0.02     0.551    0.018     0.031    0.047   0.008
7           0.043  278       0.02     0.548    0.018     0.030    0.047   0.007
8           0.014  295       0.02     0.533    0.008     0.017    0.029   0.005
9           0.029  347       0.02     0.492    0.014     0.024    0.038   0.006
10          0.037  349       0.02     0.491    0.017     0.029    0.043   0.007
11          0.039  358       0.02     0.484    0.018     0.030    0.045   0.007
12          0.018  396       0.02     0.459    0.010     0.019    0.030   0.005
13          0.028  431       0.02     0.438    0.015     0.024    0.037   0.006
14          0.025  441       0.02     0.433    0.013     0.023    0.035   0.005
15          0.027  477       0.02     0.414    0.015     0.024    0.036   0.006
16          0.045  484       0.02     0.410    0.023     0.035    0.050   0.007
17          0.030  494       0.02     0.405    0.016     0.026    0.039   0.006
18          0.022  501       0.02     0.402    0.012     0.021    0.032   0.005
19          0.028  505       0.02     0.400    0.015     0.025    0.036   0.005
20          0.020  540       0.02     0.384    0.012     0.020    0.031   0.005
21          0.028  563       0.02     0.374    0.016     0.025    0.037   0.005
22          0.024  593       0.02     0.362    0.014     0.022    0.033   0.005
23          0.015  602       0.02     0.358    0.010     0.017    0.026   0.004
24          0.024  629       0.02     0.348    0.014     0.023    0.033   0.005
25          0.020  636       0.02     0.346    0.012     0.020    0.030   0.005
26          0.048  729       0.02     0.316    0.027     0.039    0.053   0.007
27          0.031  849       0.02     0.284    0.019     0.028    0.038   0.005
28          0.027  914       0.02     0.269    0.017     0.025    0.035   0.005
29          0.021  940       0.02     0.264    0.014     0.021    0.030   0.004
30          0.029 1193       0.02     0.220    0.020     0.027    0.036   0.004
31          0.020 1340       0.02     0.201    0.014     0.020    0.027   0.003
colMeans    0.029  517       0.02     0.438    0.015     0.025    0.037   0.006
\end{CodeOutput}
\end{CodeChunk}
For reference, we need to type `\code{R> print(p, sort = FALSE)}' instead of `\code{R> p}' in order to list hospitals by the order of data input in the above output. `\code{R> p}' automatically sorts the output by the increasing order of caseload ($n_{j}$).

The output contains information about sample mean (\code{obs.mean}), caseload (\code{n}), known prior mean ($\lambda_{0}$), shrinkage estimate ($\hat{B}_{j}$), lower interval, posterior mean ($E(\lambda_{j}\vert \textbf{z})$), upper interval, and standard deviation of posterior distribution ($sd(\lambda_{j}\vert \textbf{z})$).

As we can see in (6), the posterior mean, $(1-B_{j})y_{j} + B_{j}\lambda_{0}$, is a linear function of shrinkage, $B_{j}\equiv r / (r + n_{j})$, locating between the sample mean and prior mean ($\lambda_{0}=0.02$). It makes sense because $r$ can be interpreted as the amount of prior information and $n_{j}$ as the amount of observed information. If the second level has more information than the first level, then the sample mean shrinks towards the prior mean more than 50\%. This point is clear in the above output because, as caseload increases, shrinkage decreases, depending less on the prior distribution.

@@

  post.mode.alpha post.sd.alpha
1          -5.818         0.411
\end{CodeOutput}
\end{CodeChunk}
Since estimated $\alpha$ is -5.818, we can easily calculate $\hat{r}=\textrm{exp}(5.818)=336$, which is a good indicator of how valuable and informative the hypothetical second-level hierarchy is. As mentioned before, $r$ can be interpreted as the amount of prior information, although we do not know its true value. So, \code{gbp} asked data how much credit they wanted to give to this hypothetical second-level hierarchy. $\hat{r}$ = 336 is the answer from the data. It means that observed sample means of hospitals whose caseloads are less than 336 will shrink toward the prior mean (0.020) more than 50\%.

For example, the shrinkage estimate of the first hospital ($\hat{B}_{1}= 0.834$) was calculated by 336 / (336 + 67), where 67 is its caseload ($n_{1}$), and its posterior mean is $(1-0.834)*0.045 + 0.834 * 0.020=0.024$. As for this hospital, using more information from the prior distribution is an appropriate choice because its observed amount of information (67) is far less than the amount of state-level information (336).

@@

The regression towards the mean (RTTM) is obvious in the left-side graph; the observed sample means, empty dots on the upper line, are shrinking towards the known second-level mean (a blue vertical line at 0.02) to the different extents. Note that some hospitals' ranks have changed by shrinking much harder towards 0.02 than others. For example, the empty square symbol at the crossing point of the two left-most lines (8th and 23rd hospitals on the list above) indicates that seemingly safest hospital among 31 hospitals in terms of the observed sample mean was not actually safer than the second safest hospital. Without this hierarchical modeling, we might have made a wrong decision in choosing a hospital.
